# Inorganic Phosphorus

## Contributors
## Overview

## Model Description

This module supports one state variable to capture the dissolved phosphate concentration, $PO_4$ (also referred to as *Filterable Reactive Phosphorus*, $FRP$), and optionally users can account for sorbed phosphate, denoted $PO_4^{ads}$.
The module is a low-level module that supports the processes shaping inorganic phosphorus dynamics and is designed to be linked to by other modules that interact with phosphate. The dynamics of $PO_4$ are summarised as:

\begin{eqnarray}
\frac{D}{Dt}PO_4 =  \color{darkgray}{ \mathbb{M} + \mathcal{S} } \quad
&+&   \overbrace{\check{f}_{wetdep}^{frp}+\hat{f}_{sed}^{frp} - f_{sorp}^{frp} }^\text{aed_phosphorus} \\ (\#eq:phs1)
&+& \color{brown}{ f_{min}^{DOP} - f_{gpp}^{PHY}- f_{gpp}^{MAG}  } \\ \nonumber
\end{eqnarray}

where $\mathbb{M}$ and $\mathcal{S}$ refer to water mixing and boundary source terms, respectively, and the coloured $\color{brown}{f}$ terms reflect the optionally configurable contributions from other modules; these include the breakdown of $DOP$ by aerobic heterotrophic bacteria to $PO_4$, and photosynthetic phosphate uptake by phytoplankton and macroalgae.

The sorbed $PO_4$ pool is resolved as:

\begin{eqnarray}
\frac{D}{Dt}PO_4^{ads} =  \color{darkgray}{ \mathbb{M} + \mathcal{S} } \quad
&+&   \overbrace{\check{f}_{drydep}^{pip}+\hat{f}_{resus}^{pip} + f_{sorp}^{frp} - f_{set}^{pip} }^\text{aed_phosphorus} \\ (\#eq:phs2)
\end{eqnarray}


###	Process Descriptions

#### Phosphate sorption - desorption

#### Atmospheric deposition

#### Dissolved sediment flux

#### Sedimentation and resuspension



###	Optional Module Links

Other phosphate sources/sinks are indicated in Eq \@ref(eq:phs1) for $PO_4$. These include:

-	aed_oxygen: Oxygen is used to control the rate of sediment release of phosphate.
-	aed_carbon: pH can used to influence the sorption dynamics of phosphate.
-	aed_organic_matter: organic matter minerization replenishes $PO_4$.
-	aed_phytoplankton: phytoplankton photosynthesis consumes $PO_4$.
-	aed_geochemistry: optional link to simulate the phosphate precipitation.
-	aed_sedflux: modules to update the sediment-water exchange rate of $PO_4$.

###	Feedbacks to the Host Model

The inorganic phosphorus module has no feedbacks to the host hydrodynamic model.

###	Variable Summary

The default variables created by this module, and the optionally required linked variables needed for full functionality of this module are summarised in Table \@ref(tab:13-statetable). The diagnostic outputs able to be output are summarised in Table \@ref(tab:13-diagtable).

#### State variables {-}


```{r 13-statetable, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(readxl)
library(rmarkdown)
theSheet <- read_excel('tables/aed_variable_tables.xlsx', sheet = 8)
theSheet <- theSheet[theSheet$Table == "State variable",]
theSheetGroups <- unique(theSheet$Group)
theSheet$`AED name` <- paste0("`",theSheet$`AED name`,"`")

for(i in seq_along(theSheet$Symbol)){
  if(!is.na(theSheet$Symbol[i])==TRUE){
    theSheet$Symbol[i] <- paste0("$$",theSheet$Symbol[i],"$$")
  } else {
    theSheet$Symbol[i] <- NA
  }
}
for(i in seq_along(theSheet$Unit)){
  if(!is.na(theSheet$Unit[i])==TRUE){
    theSheet$Unit[i] <- paste0("$$\\small{",theSheet$Unit[i],"}$$")
  } else {
    theSheet$Unit[i] <- NA
  }
}
for(i in seq_along(theSheet$Comments)){
  if(!is.na(theSheet$Comments[i])==TRUE){
    theSheet$Comments[i] <- paste0("",theSheet$Comments[i],"")
  } else {
    theSheet$Comments[i] <- " "
  }
}

kbl(theSheet[,3:NCOL(theSheet)], caption = "Phosphorus - *state* variables", align = "l",) %>%
  pack_rows(theSheetGroups[1],
            min(which(theSheet$Group == theSheetGroups[1])),
            max(which(theSheet$Group == theSheetGroups[1])),
            background = '#ebebeb') %>%
  pack_rows(theSheetGroups[2],
            min(which(theSheet$Group == theSheetGroups[2])),
            max(which(theSheet$Group == theSheetGroups[2])),
            background = '#ebebeb') %>%
  row_spec(0, background = "#14759e", bold = TRUE, color = "white") %>%
	kable_styling(full_width = F,font_size = 11) %>%
	column_spec(2, width_min = "7em") %>%
	column_spec(3, width_max = "18em") %>%
	column_spec(4, width_min = "10em") %>%
	column_spec(5, width_min = "5em") %>%
	column_spec(7, width_min = "10em") %>%
	column_spec(7, width_max = "20em") %>%
  scroll_box(width = "770px", height = "465px",
             fixed_thead = FALSE)
```



#### Diagnostics{-}



```{r 13-diagtable, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(readxl)
library(rmarkdown)
theSheet <- read_excel('tables/aed_variable_tables.xlsx', sheet = 8)
theSheet <- theSheet[theSheet$Table == "Diagnostics",]
theSheetGroups <- unique(theSheet$Group)
theSheet$`AED name` <- paste0("`",theSheet$`AED name`,"`")

for(i in seq_along(theSheet$Symbol)){
  if(!is.na(theSheet$Symbol[i])==TRUE){
    theSheet$Symbol[i] <- paste0("$$",theSheet$Symbol[i],"$$")
  } else {
    theSheet$Symbol[i] <- NA
  }
}
for(i in seq_along(theSheet$Unit)){
  if(!is.na(theSheet$Unit[i])==TRUE){
    theSheet$Unit[i] <- paste0("$$\\small{",theSheet$Unit[i],"}$$")
  } else {
    theSheet$Unit[i] <- NA
  }
}
for(i in seq_along(theSheet$Comments)){
  if(!is.na(theSheet$Comments[i])==TRUE){
    theSheet$Comments[i] <- paste0("",theSheet$Comments[i],"")
  } else {
    theSheet$Comments[i] <- " "
  }
}

kbl(theSheet[,3:NCOL(theSheet)], caption = "Phosphorus - *diagnostic* variables", align = "l",) %>%
  pack_rows(theSheetGroups[1],
            min(which(theSheet$Group == theSheetGroups[1])),
            max(which(theSheet$Group == theSheetGroups[1])),
            background = '#ebebeb') %>%
  pack_rows(theSheetGroups[2],
            min(which(theSheet$Group == theSheetGroups[2])),
            max(which(theSheet$Group == theSheetGroups[2])),
            background = '#ebebeb') %>%
  row_spec(0, background = "#14759e", bold = TRUE, color = "white") %>%
  kable_styling(full_width = F,font_size = 11) %>%
            column_spec(2, width_min = "7em") %>%
	          column_spec(3, width_max = "18em") %>%
	          column_spec(4, width_min = "10em") %>%
	          column_spec(5, width_min = "6em") %>%
						column_spec(7, width_min = "10em") %>%
	          scroll_box(width = "770px", height = "245px",
            fixed_thead = FALSE)
```

<br>


###	Parameter Summary

The parameters and settings used by this module are summarised in Table \@ref(tab:13-parstable).


```{r 13-parstable, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(readxl)
library(rmarkdown)
theSheet <- read_excel('tables/aed_variable_tables.xlsx', sheet = 8)
theSheet <- theSheet[theSheet$Table == "Parameter",]
theSheetGroups <- unique(theSheet$Group)
theSheet$`AED name` <- paste0("`",theSheet$`AED name`,"`")

for(i in seq_along(theSheet$Symbol)){
  if(!is.na(theSheet$Symbol[i])==TRUE){
    theSheet$Symbol[i] <- paste0("$$",theSheet$Symbol[i],"$$")
  } else {
    theSheet$Symbol[i] <- " "
  }
}
for(i in seq_along(theSheet$Unit)){
  if(!is.na(theSheet$Unit[i])==TRUE){
    theSheet$Unit[i] <- paste0("$$\\small{",theSheet$Unit[i],"}$$")
  } else {
    theSheet$Unit[i] <- NA
  }
}
for(i in seq_along(theSheet$Comments)){
  if(!is.na(theSheet$Comments[i])==TRUE){
    theSheet$Comments[i] <- paste0("",theSheet$Comments[i],"")
  } else {
    theSheet$Comments[i] <- "-"
  }
}

kbl(theSheet[,3:NCOL(theSheet)], caption = "Phosphorus module parameters and configuration options", align = "l",) %>%
  pack_rows(theSheetGroups[1],
            min(which(theSheet$Group == theSheetGroups[1])),
            max(which(theSheet$Group == theSheetGroups[1])),
            background = '#ebebeb') %>%
  pack_rows(theSheetGroups[2],
            min(which(theSheet$Group == theSheetGroups[2])),
            max(which(theSheet$Group == theSheetGroups[2])),
            background = '#ebebeb') %>%
  pack_rows(theSheetGroups[3],
					  min(which(theSheet$Group == theSheetGroups[3])),
					  max(which(theSheet$Group == theSheetGroups[3])),
					  background = '#ebebeb') %>%
  pack_rows(theSheetGroups[4],
					  min(which(theSheet$Group == theSheetGroups[4])),
					  max(which(theSheet$Group == theSheetGroups[4])),
					  background = '#ebebeb') %>%
  row_spec(0, background = "#14759e", bold = TRUE, color = "white") %>%
  kable_styling(full_width = F,font_size = 11) %>%
	column_spec(2, width_min = "7em") %>%
	column_spec(3, width_max = "19em") %>%
	column_spec(4, width_min = "10em") %>%
	column_spec(5, width_min = "5em") %>%
	column_spec(7, width_min = "10em") %>%
  scroll_box(width = "770px", height = "1300px",
             fixed_thead = FALSE)
```
<br>

## Setup & Configuration


An example `aed.nml` parameter specification block for the `aed_phosphorus` module that is modelling dissolved and particulare (sorbed) $PO_4$ is shown below:
```{fortran, eval = FALSE}
  &aed_phosphorus
    !-- Initial value
     frp_initial          = 0.25
    !-- Sediment flux
     Fsed_frp             = 0.08
     Ksed_frp             = 30.0
     theta_sed_frp        = 1.08
     phosphorus_reactant_variable ='OXY_oxy'
    !-- PO4 adsorption
     simPO4Adsorption     = .true.
     ads_use_external_tss = .true.
     PO4AdsorptionModel   = 1
     Kpo4p                = 0.01
     w_po4ads             = -0.01
  /
```

<br>
An example `aed.nml` parameter specification block for the `aed_phosphorus` module that includes all options is shown below:
```{fortran, eval = FALSE}
  &aed_phosphorus
    !-- Initial value
     frp_initial = 0.25
    !-- Sediment flux
     Fsed_frp = 0.08
     Ksed_frp = 30.0
     theta_sed_frp = 1.08
     phosphorus_reactant_variable ='OXY_oxy'
     Fsed_frp_variable ='SDF_Fsed_frp'
    !-- PO4 adsorption
     simPO4Adsorption = .true.
     ads_use_external_tss = .false.
     po4sorption_target_variable ='NCS_ss1'
     PO4AdsorptionModel = 2
     Kpo4p = 0.01
     Kadsratio= 1.0
     Qmax = 1.0
     w_po4ads = -9999     ! Note: -9999 links PO4-ad settling to target_variable
     ads_use_pH = .false.
     pH_variable = 'CAR_pH'
    !-- Atmospheric deposition
     simDryDeposition = .false.
     atm_pip_dd = 0.00
     simWetDeposition = .true.
     atm_frp_conc = 0.01
  /
```

## Case Studies & Examples
###	Case Study
###	Publications
