# Silica

## Contributors
## Overview
## Model Description
###	Process Descriptions
###	Optional Module Links

###	Feedbacks to the Host Model

The silica module has no feedbacks to the host hydrodynamic model.

###	Variable Summary

The default variables created by this module, and the optionally required linked variables needed for full functionality of this module are summarised in Table \@ref(tab:14-statetable). The diagnostic outputs able to be output are summarised in Table \@ref(tab:14-diagtable).

#### State variables {-}

```{r 14-statetable, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(readxl)
library(rmarkdown)
theSheet <- read_excel('tables/aed_variable_tables.xlsx', sheet = 9)
theSheet <- theSheet[theSheet$Table == "State variable",]
theSheetGroups <- unique(theSheet$Group)
theSheet$`AED name` <- paste0("`",theSheet$`AED name`,"`")

for(i in seq_along(theSheet$Symbol)){
  if(!is.na(theSheet$Symbol[i])==TRUE){
    theSheet$Symbol[i] <- paste0("$$",theSheet$Symbol[i],"$$")
  } else {
    theSheet$Symbol[i] <- NA
  }
}
for(i in seq_along(theSheet$Unit)){
  if(!is.na(theSheet$Unit[i])==TRUE){
    theSheet$Unit[i] <- paste0("$$\\small{",theSheet$Unit[i],"}$$")
  } else {
    theSheet$Unit[i] <- NA
  }
}
for(i in seq_along(theSheet$Comments)){
  if(!is.na(theSheet$Comments[i])==TRUE){
    theSheet$Comments[i] <- paste0("",theSheet$Comments[i],"")
  } else {
    theSheet$Comments[i] <- " "
  }
}

kbl(theSheet[,3:NCOL(theSheet)], caption = "Silica - *state* variables", align = "l",) %>%
  pack_rows(theSheetGroups[1],
            min(which(theSheet$Group == theSheetGroups[1])),
            max(which(theSheet$Group == theSheetGroups[1])),
            background = '#ebebeb') %>%
  pack_rows(theSheetGroups[2],
            min(which(theSheet$Group == theSheetGroups[2])),
            max(which(theSheet$Group == theSheetGroups[2])),
            background = '#ebebeb') %>%
  row_spec(0, background = "#14759e", bold = TRUE, color = "white") %>%
	kable_styling(full_width = F,font_size = 11) %>%
	column_spec(2, width_min = "7em") %>%
	column_spec(3, width_max = "18em") %>%
	column_spec(4, width_min = "10em") %>%
	column_spec(5, width_min = "5em") %>%
	column_spec(7, width_min = "10em") %>%
	column_spec(7, width_max = "20em") %>%
  scroll_box(width = "770px", height = "200px",
             fixed_thead = FALSE)
```

#### Diagnostics{-}

```{r 14-diagtable, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(readxl)
library(rmarkdown)
theSheet <- read_excel('tables/aed_variable_tables.xlsx', sheet = 9)
theSheet <- theSheet[theSheet$Table == "Diagnostics",]
theSheetGroups <- unique(theSheet$Group)
theSheet$`AED name` <- paste0("`",theSheet$`AED name`,"`")

for(i in seq_along(theSheet$Symbol)){
  if(!is.na(theSheet$Symbol[i])==TRUE){
    theSheet$Symbol[i] <- paste0("$$",theSheet$Symbol[i],"$$")
  } else {
    theSheet$Symbol[i] <- NA
  }
}
for(i in seq_along(theSheet$Unit)){
  if(!is.na(theSheet$Unit[i])==TRUE){
    theSheet$Unit[i] <- paste0("$$\\small{",theSheet$Unit[i],"}$$")
  } else {
    theSheet$Unit[i] <- NA
  }
}
for(i in seq_along(theSheet$Comments)){
  if(!is.na(theSheet$Comments[i])==TRUE){
    theSheet$Comments[i] <- paste0("",theSheet$Comments[i],"")
  } else {
    theSheet$Comments[i] <- " "
  }
}

kbl(theSheet[,3:NCOL(theSheet)], caption = "Silica - *diagnostic* variables", align = "l",) %>%
  pack_rows(theSheetGroups[1],
            min(which(theSheet$Group == theSheetGroups[1])),
            max(which(theSheet$Group == theSheetGroups[1])),
            background = '#ebebeb') %>%
  row_spec(0, background = "#14759e", bold = TRUE, color = "white") %>%
  kable_styling(full_width = F,font_size = 11) %>%
            column_spec(2, width_min = "7em") %>%
	          column_spec(3, width_max = "18em") %>%
	          column_spec(4, width_min = "10em") %>%
	          column_spec(5, width_min = "6em") %>%
						column_spec(7, width_min = "10em") %>%
	          scroll_box(width = "770px", height = "120px",
            fixed_thead = FALSE)
```

<br>


###	Parameter Summary

The parameters and settings used by this module are summarised in Table \@ref(tab:14-parstable).


```{r 14-parstable, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(readxl)
library(rmarkdown)
theSheet <- read_excel('tables/aed_variable_tables.xlsx', sheet = 9)
theSheet <- theSheet[theSheet$Table == "Parameter",]
theSheetGroups <- unique(theSheet$Group)
theSheet$`AED name` <- paste0("`",theSheet$`AED name`,"`")

for(i in seq_along(theSheet$Symbol)){
  if(!is.na(theSheet$Symbol[i])==TRUE){
    theSheet$Symbol[i] <- paste0("$$",theSheet$Symbol[i],"$$")
  } else {
    theSheet$Symbol[i] <- " "
  }
}
for(i in seq_along(theSheet$Unit)){
  if(!is.na(theSheet$Unit[i])==TRUE){
    theSheet$Unit[i] <- paste0("$$\\small{",theSheet$Unit[i],"}$$")
  } else {
    theSheet$Unit[i] <- NA
  }
}
for(i in seq_along(theSheet$Comments)){
  if(!is.na(theSheet$Comments[i])==TRUE){
    theSheet$Comments[i] <- paste0("",theSheet$Comments[i],"")
  } else {
    theSheet$Comments[i] <- "-"
  }
}

kbl(theSheet[,3:NCOL(theSheet)], caption = "Silica module parameters and configuration options", align = "l",) %>%
  pack_rows(theSheetGroups[1],
            min(which(theSheet$Group == theSheetGroups[1])),
            max(which(theSheet$Group == theSheetGroups[1])),
            background = '#ebebeb') %>%
  pack_rows(theSheetGroups[2],
            min(which(theSheet$Group == theSheetGroups[2])),
            max(which(theSheet$Group == theSheetGroups[2])),
            background = '#ebebeb') %>%
  row_spec(0, background = "#14759e", bold = TRUE, color = "white") %>%
  kable_styling(full_width = F,font_size = 11) %>%
	column_spec(2, width_min = "7em") %>%
	column_spec(3, width_max = "19em") %>%
	column_spec(4, width_min = "10em") %>%
	column_spec(5, width_min = "5em") %>%
	column_spec(7, width_min = "10em") %>%
  scroll_box(width = "770px", height = "500px",
             fixed_thead = FALSE)
```
<br>

## Setup & Configuration

An example `aed.nml` parameter specification block for the `aed_silica` module is shown below:
```{fortran, eval = FALSE}
  &aed_silica
    !-- Initial value
     rsi_initial              = 12.5
    !-- Sediment flux
     Fsed_rsi                 = 1.8
     Ksed_rsi                 = 50.0
     theta_sed_rsi            = 1.08
     silica_reactant_variable = 'OXY_oxy'
     Fsed_rsi_variable        = 'SDF_Fsed_rsi'
  /
```

## Case Studies & Examples
###	Case Study
###	Publications
