# (PART)  Supporting Material {-}

# Generic Utilities & Functions {#genericUtilitiesFunctions}

<!-- ## Overview -->


<!-- ## Physico-chemical -->

### Gas Transfer {#GAS}

This section descrives some popularly used gas exchange velocity that are optionally selcted in various AED modules. The key factors of wind speed, current speed, and water depths are considered.

When choosing an algorithm, consider the following selection criteria:

1. Environment: ocean; lake; estuary/river;
2. Study for which gas: O~2~, CH~4~, N~2~O, CO~2~
3. Variables: wind speeds ($U$), current speeds ($v$), depths ($h$)
4. Schmidt number normalization: yes/no. $Sc = \frac{\mu }{d} = \frac{\text{kinematic viscosity}}{\text{diffusion}}$

<br>

<center>
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
table1 <- read.csv("tables/generic_utilities_functions-1.csv", check.names=FALSE)
kable(table1,"html", escape = F, align = "c", caption = "Options for schmidt number calculation in gas-exchange calculations.",
      bootstrap_options = "hover") %>%
  kable_styling(table1, bootstrap_options = "hover",
                full_width = F, position = "center",
                font_size = 12) %>%
  column_spec(1, width_min = "6em") %>%
  column_spec(2, width_min = "6em") %>%
  column_spec(3, width_min = "6em") %>%
  row_spec(1:6, background = 'white') %>%
  scroll_box(width = "770px", height = "380px",
  fixed_thead = FALSE)
```
</center>
<br>
<center>
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
table2 <- read.csv("tables/generic_utilities_functions-2.csv", check.names=FALSE)
kable(table2,"html", escape = F, align = "c", caption = "Options for gas exchnage velocity calculation.",
      bootstrap_options = "hover") %>%
  kable_styling(table2, bootstrap_options = "hover",
                full_width = F, position = "center",
                font_size = 12) %>%
  column_spec(1:7, width_min = "6em") %>%
  row_spec(1:8, background = 'white') %>%
  scroll_box(width = "device-width", height = "900px",
  fixed_thead = FALSE)
```
</center>

<!-- ## Biological -->
