# (PART)  Supporting Material {-} 

# Generic Utilities & Functions

## Overview


## Physico-chemical

### Gas Transfer

Target: review and select a few most popularly used gas exchange velocity to be used in AED modules. Note only key factors of wind speed, current speed, and water depths are considered. 

Selection criteria:

1. Environment: ocean; lake; estuary/river;
2. Study for which gas: O~2~, CH~4~, N~2~O, CO~2~
3. Variables: wind speeds ($U$), current speeds ($v$), depths ($h$)
4. Schmidt number normalization: yes/no. $Sc = \frac{\mu }{d} = \frac{\text{kinematic viscosity}}{\text{diffusion}}$


<center>
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
table1 <- read.csv("tables/generic_utilities_functions-1.csv", check.names=FALSE)
kable(table1,"html", escape = F, align = "c") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position = "center") %>%
  column_spec(1, width_min = "12em") %>%
  column_spec(2, width_min = "17em") %>%
  column_spec(3, width_min = "10em") %>%
  scroll_box(width = "725px", height = "400px",
  fixed_thead = FALSE) 
```
</center>
<br>
<center>
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
table2 <- read.csv("tables/generic_utilities_functions-2.csv", check.names=FALSE)
kable(table2,"html", escape = F, align = "c") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position = "center") %>%
  column_spec(1, width_min = "12em") %>%
  column_spec(2, width_min = "22em") %>%
  column_spec(7, width_min = "15em") %>%
  scroll_box(width = "725px", height = "500px",
  fixed_thead = FALSE) 
```
</center>

## Biological



